### 2017-10-22 paper/sections/0_abstract.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Data center scale application developers are abandoning POSIX because the file
system metadata synchronization and serialization overheads of providing strong
consistency and durability are too costly -- and often unnecessary -- for their
applications.  Unfortunately, designing near-POSIX file systems excludes
applications that rely on strong consistency or durability, forcing developers
to re-write their applications or deploy them on a different system. Mounting
multiple systems in the global namespace forces users to (1) provision separate
storage clusters and (2) manually move data across systems.  We present a
framework and API that lets clients dynamically specify their
consistency/durability requirements and assign them to subtrees in the
namespace, allowing users to optimize subtrees within the same namespace for
different workloads.  We confirm the performance benefits of techniques
presented in related work but also explore new consistency/durability metadata
designs, all over the same storage system.  By custom fitting a subtree to a
create-heavy application, we show an 8\(\times\) speedup and scale further
(twice as many servers) than the system we implemented our prototype on.

>>>>

Data center scale application developers are abandoning POSIX IO because the file
system metadata synchronization and serialization overheads of providing strong
consistency and durability are too costly -- and often unnecessary -- for their
applications.  Unfortunately, designing near-POSIX IO file systems excludes
applications that rely on strong consistency or durability, forcing developers
to re-write their applications or deploy them on a different system. Mounting
multiple systems in the global namespace forces users to (1) provision separate
storage clusters and (2) manually move data across systems.  We present a
framework and API that lets clients dynamically specify their
consistency/durability requirements and assign them to subtrees in the
namespace, allowing users to optimize subtrees within the same namespace for
different workloads.  We confirm the performance benefits of techniques
presented in related work but also explore new consistency/durability metadata
designs, all over the same storage system.  By custom fitting a subtree to a
create-heavy application, we show an 8\(\times\) speedup and scale further
(twice as many servers) than the system we implemented our prototype on.

### 2017-10-22 paper/sections/1_introduction.tex
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many batch style jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N and N:1 create
patterns~\cite{bent_plfs_2009}).

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{ Illustration of subtrees with different semantics co-existing in a
global namespace.  For performance, clients can relax consistency on their
subtree (HDFS) or even decouple the subtree and move it locally (BatchFS,
RAMDisk). Decoupled subtrees can further relax durability for even better
performance.  Clients that require stronger guarantees (POSIX) can still reside
in the same namespace.  }\label{fig:subtree-policies}
\end{figure}

@@

% What did we do
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a potential
setup in our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art
architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
paper, although the HDFS semantics and their performance is explored
in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
consistency because it does not guarantee that all updates are immediately seen
by all servers; the BatchFS and RAMDisk subtrees are decoupled from the global
namespace and have similar consistency/durability behavior to those systems;
and the POSIX subtree retains the rigidity of POSIX's strong consistency.
Subtrees without policies inherit the consistency/durability semantics of the
parent and future work will examine embeddable or composable policies.

@@

In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
cost of POSIX consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes the Cudele
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places Cudele in the context of other related
work.

>>>>

% What is HPC doing?
To address this, developers are relaxing the consistency and durability
semantics in the file system because weaker guarantees are sufficient for their
applications. For example, many batch style jobs do not need the strong
consistency that the file system provides, so
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
do more client-side processing and merge updates when the job is done. HPC
developers are turning to these non-POSIX IO solutions because their applications
are well-understood ({\it e.g.}, well-defined read/write phases,
synchronization only needed during certain phases, workflows describing
computation, etc.) and because these applications wreak havoc on file systems designed for
general-purpose workloads ({\it e.g.}, checkpoint-restart's N:N and N:1 create
patterns~\cite{bent_plfs_2009}).

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{ Illustration of subtrees with different semantics co-existing in a
global namespace.  For performance, clients can relax consistency on their
subtree (HDFS) or even decouple the subtree and move it locally (BatchFS,
RAMDisk). Decoupled subtrees can further relax durability for even better
performance.  Clients that require stronger guarantees (POSIX IO) can still reside
in the same namespace.  }\label{fig:subtree-policies}
\end{figure}

@@

% What did we do
To address this problem, we present an API and framework that lets developers
dynamically control the consistency and durability guarantees for subtrees in
the file system namespace.  Figure~\ref{fig:subtree-policies} shows a potential
setup in our proposed system where a single global namespace has subtrees for
applications optimized with techniques from different state-of-the-art
architectures.  The HDFS\footnote{HDFS itself is not directly evaluated in this
paper, although the HDFS semantics and their performance is explored
in~\S\ref{sec:use-case-1}} subtree has weaker consistency than strong
consistency because it does not guarantee that all updates are immediately seen
by all servers; the BatchFS and RAMDisk subtrees are decoupled from the global
namespace and have similar consistency/durability behavior to those systems;
and the POSIX IO subtree retains the rigidity of POSIX IO's strong consistency.
Subtrees without policies inherit the consistency/durability semantics of the
parent and future work will examine embeddable or inheritable policies.

@@

In the remainder of the paper, Section~\ref{sec:posix-overheads} quantifies the
cost of POSIX IO consistency and system-defined durability and
Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
prototype and API. Section~\ref{sec:implementation} describes the Cudele
mechanisms and shows how re-using internal subsystems results in an
implementation of less than 500 lines of code. The evaluation in
Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
explored and previously unexplored metadata designs.
Section~\ref{sec:related-work} places Cudele in the context of other related
work.

### 2017-10-22 paper/sections/2_related_work.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

The bottlenecks associated with accessing POSIX file system metadata are not limited

@@

%consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates

@@

%\subsection{Relaxing POSIX}

% POSIX 
POSIX workloads require strong consistency and many file systems improve
performance by reducing the number of remote calls per operation ({\it i.e.}
RPC amplification). As discussed in the previous section, caching with leases and
replication are popular approaches to reducing the overheads of path traversals
but their performance is subject to cache locality and the amount of available
resources, respectively; for random workloads larger than the cache extra RPCs
hurt performance~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} and for write heavy workloads with more
resources the RPCs for invalidations are harmful. Another approach to reducing
RPCs is to use leases or capabilities.

@@

% Non POSIX
High performance computing has unique requirements for file systems ({\it
e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
make relaxing POSIX sensible.  BatchFS assumes the application coordinates
accesses to the namespace, so the clients can batch local operations and merge
with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
traffic using subtree snapshots for non-conflicting workloads and middleware
for conflicting workloads. MarFS gives users the ability to lock
``project directories" and allocate GPFS clusters for demanding metadata
workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
tier; applications lock the namespace so that metadata can be accessed more quickly.
Unfortunately, decoupling the namespace has costs: (1) merging metadata state
back into the global namespace is slow; (2) failures are local to the failing
node; and (3) the systems are not backwards compatible.

% NON POSIX
%Decoupling the namespaces has many advantages, including improved scalability,
%higher resource utilization, and better performance.  BatchFS and DeltaFS are
%near-POSIX filesystems that give clients the ability to decouple subtrees from
%the namespace so that the applications can execute metadata operations without
%synchronization and serialization.  These operations are applied to a local
%snapshot of the file system namespace and conflicts are resolved either by the
%application or by an external service . Applications link into a metadata
%server library to reduce resource utilization and code paths (e.g., no daemons
%and less interprocess communication).

@@

For (3), decoupled namespace approaches sacrifice POSIX going as far as
requiring the application to link against the systems they want to talk to. In
today's world of software defined caching, this can be a problem for large data
centers with many types and tiers of storage. Despite well-known performance
problems POSIX and REST are the dominant APIs for data transfer.

>>>>

The bottlenecks associated with accessing POSIX IO file system metadata are not limited

@@

%consistency ({e.g.}, POSIX IO).  For example, ShardFS pessimistically replicates

@@

%\subsection{Relaxing POSIX IO}

% POSIX IO 
POSIX IO workloads require strong consistency and many file systems improve
performance by reducing the number of remote calls per operation ({\it i.e.}
RPC amplification). As discussed in the previous section, caching with leases and
replication are popular approaches to reducing the overheads of path traversals
but their performance is subject to cache locality and the amount of available
resources, respectively; for random workloads larger than the cache extra RPCs
hurt performance~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} and for write heavy workloads with more
resources the RPCs for invalidations are harmful. Another approach to reducing
RPCs is to use leases or capabilities.

@@

% Non POSIX IO
High performance computing has unique requirements for file systems ({\it
e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
make relaxing POSIX IO sensible.  BatchFS assumes the application coordinates
accesses to the namespace, so the clients can batch local operations and merge
with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
traffic using subtree snapshots for non-conflicting workloads and middleware
for conflicting workloads. MarFS gives users the ability to lock
``project directories" and allocate GPFS clusters for demanding metadata
workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
tier; applications lock the namespace so that metadata can be accessed more quickly.
Unfortunately, decoupling the namespace has costs: (1) merging metadata state
back into the global namespace is slow; (2) failures are local to the failing
node; and (3) the systems are not backwards compatible.

% NON POSIX IO
%Decoupling the namespaces has many advantages, including improved scalability,
%higher resource utilization, and better performance.  BatchFS and DeltaFS are
%near-POSIX IO filesystems that give clients the ability to decouple subtrees from
%the namespace so that the applications can execute metadata operations without
%synchronization and serialization.  These operations are applied to a local
%snapshot of the file system namespace and conflicts are resolved either by the
%application or by an external service . Applications link into a metadata
%server library to reduce resource utilization and code paths (e.g., no daemons
%and less interprocess communication).

@@

For (3), decoupled namespace approaches sacrifice POSIX IO going as far as
requiring the application to link against the systems they want to talk to. In
today's world of software defined caching, this can be a problem for large data
centers with many types and tiers of storage. Despite well-known performance
problems POSIX IO and REST are the dominant APIs for data transfer.

### 2017-10-22 paper/sections/3_posix_overheads.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{POSIX Overheads}
\label{sec:posix-overheads}

@@

% CITEME PDSW paper
In our examination of the overheads of POSIX we benchmark and analyze CephFS,
the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
data and metadata. To show how the file system behaves under high metadata load
we use a create-heavy workload. Create-heavy workloads are studied the most in
HPC research because of the checkpoint/restart use case but they also happen to
stress the underlying storage system the most.  During this process we
discovered, based on the analysis and breakdown of costs, that durability and
consistency have high overhead but we urge the read to keep in mind that this
file system is an implementation of one set of design decisions and our goal
here is to highlight the effect that those decisions have on performance.

@@

% what is durability
While durability is not specified by POSIX, users expect that files they create
or modify survive failures.  We define three types of durability: global,
local, and none.  Global
durability means that the client or server can fail at any time and metadata
will not be lost. Local durability means that metadata can be lost if the
client or server stays down after a failure. None means that metadata is volatile
and that the system provides no guarantees when clients or servers fail. 
None is different than local durability because regardless of the type failure,
metadata will be lost when components die in a None configuration.\\

@@

Access to metadata in a POSIX-compliant file system is strongly consistent, so
reads and writes to the same inode or directory are globally ordered.  The
synchronization and serialization machinery needed to ensure that all clients
see the same state has high overhead.\\

>>>>

\section{POSIX IO Overheads}
\label{sec:posix-overheads}

@@

% CITEME PDSW paper
In our examination of the overheads of POSIX IO we benchmark and analyze CephFS,
the file system that uses Ceph's object store ({\it i.e.} RADOS) to store its
data and metadata. To show how the file system behaves under high metadata load
we use a create-heavy workload. Create-heavy workloads are studied the most in
HPC research because of the checkpoint/restart use case but they also happen to
stress the underlying storage system the most.  During this process we
discovered, based on the analysis and breakdown of costs, that durability and
consistency have high overhead but we urge the read to keep in mind that this
file system is an implementation of one set of design decisions and our goal
here is to highlight the effect that those decisions have on performance.

@@

% what is durability
While durability is not specified by POSIX IO, users expect that files they create
or modify survive failures.  We define three types of durability: global,
local, and none.  Global
durability means that the client or server can fail at any time and metadata
will not be lost. Local durability means that metadata can be lost if the
client or server stays down after a failure. None means that metadata is volatile
and that the system provides no guarantees when clients or servers fail. 
None is different than local durability because regardless of the type failure,
metadata will be lost when components die in a None configuration.\\

@@

Access to metadata in a POSIX IO-compliant file system is strongly consistent, so
reads and writes to the same inode or directory are globally ordered.  The
synchronization and serialization machinery needed to ensure that all clients
see the same state has high overhead.\\

### 2017-10-22 paper/sections/6_results.tex
paper: add patrick to authors and switch to POSIX IO
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
%% (5-7x) vs. (90-104x) slower
%% Does not scale?
%% Strong Consistency > 10X slower than Weak Consistency 
%{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
%decoupled namespace strategy and shows up to a \(20\times\) speedup over the
%traditional namespaces that use RPCs. Compared to the baseline the slowdown is
%\(5-7\times\) for Strong Consistency, which emulates BatchFS and
%\(90-104\times\) for Weak Consistency, which emulates DeltaFS.

@@

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

@@

Clients create files in private directories and a separate client, which we
call a ``rogue" client, creates files in each directory. This introduces false
sharing and the metadata server revokes capabilities on directories touched by
the rogue client. While HPC tries to avoid these situations with
worklflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it stills
happens in distributed file systems when users unintentionally access
directories in a shared file system.

@@

%\subsection{Discussion}
%
%Cudele and the experiments we present here prompt many avenues for future work.
%The first is embeddable policies, where child subtrees have specialized
%features but still maintain the guarantees  of their parent subtrees. We
%already show how applciations can compose guarantees and control their
%performance, knobs that we did not have before, but embeddable policies are a
%way to compose the policies themsleves.  For example, a RAMDisk subtree is
%POSIX-compliant but relaxes durability constraints, so it can reside under a
%POSIX subtree that is strongly consistent. These embeddable policies should be
%controlled and enforced by Cudele.
%
%Second is the opportunity for adding hardware in strategic places to speed up
%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
%performance of merging updates at the metadata server is limited by the speed
%of the disk. So applications that rely on merging many updates should consider
%equipping the metadata server with SSDs or more memory. This is a form of
%co-design that takes a ``dirty-slate" approach.
%
%Finally, performance prediction benefits from our cost quantification of each
%mechanism. Applications can use Cudele to microbenchmark their components and
%software, similar to what we did in Secion~\S\ref{sec:microbenchmarks}. Using
%those results, they can predict how much slower their system will be if they
%adopt stronger consistency or durability. This can also be a verification tool
%where performance that varies wildly from the predicted performance can be a
%red flag that something is wrong or that the bottleneck is not in the
%consistency or durability plane.

>>>>

%% DeltaFS vs. BatchFS vs. POSIX IO: decoupled is faster
%% (5-7x) vs. (90-104x) slower
%% Does not scale?
%% Strong Consistency > 10X slower than Weak Consistency 
%{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
%decoupled namespace strategy and shows up to a \(20\times\) speedup over the
%traditional namespaces that use RPCs. Compared to the baseline the slowdown is
%\(5-7\times\) for Strong Consistency, which emulates BatchFS and
%\(90-104\times\) for Weak Consistency, which emulates DeltaFS.

@@

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: weak vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: weak, local
%  - e.g., POSIX IO: strong, global
%  - e.g., PLFS: no consistency
%- durability: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

@@

Clients create files in private directories and a separate client, which we
call a ``rogue" client, creates files in each directory. This introduces false
sharing and the metadata server revokes capabilities on directories touched by
the rogue client. While HPC tries to avoid these situations with
workflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it still
happens in distributed file systems when users unintentionally access
directories in a shared file system.

@@

%\subsection{Discussion}
%
%Cudele and the experiments we present here prompt many avenues for future work.
%The first is embeddable policies, where child subtrees have specialized
%features but still maintain the guarantees  of their parent subtrees. We
%already show how applciations can compose guarantees and control their
%performance, knobs that we did not have before, but embeddable policies are a
%way to compose the policies themsleves.  For example, a RAMDisk subtree is
%POSIX IO-compliant but relaxes durability constraints, so it can reside under a
%POSIX IO subtree that is strongly consistent. These embeddable policies should be
%controlled and enforced by Cudele.
%
%Second is the opportunity for adding hardware in strategic places to speed up
%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
%performance of merging updates at the metadata server is limited by the speed
%of the disk. So applications that rely on merging many updates should consider
%equipping the metadata server with SSDs or more memory. This is a form of
%co-design that takes a ``dirty-slate" approach.
%
%Finally, performance prediction benefits from our cost quantification of each
%mechanism. Applications can use Cudele to microbenchmark their components and
%software, similar to what we did in Secion~\S\ref{sec:microbenchmarks}. Using
%those results, they can predict how much slower their system will be if they
%adopt stronger consistency or durability. This can also be a verification tool
%where performance that varies wildly from the predicted performance can be a
%red flag that something is wrong or that the bottleneck is not in the
%consistency or durability plane.
